Maximizing Ink in Symmetric Partial Edge Drawings of $k$-plane Graphs

### 2018-01-08 eurocg/paper.tex
hardness proof polishing

\newcommand{\ped}{\ensuremath{\textsc{PED}}\xspace}
\newcommand{\maxsped}{\ensuremath{\textsc{MaxSPED}}\xspace}
\newcommand{\ppsat}{\ensuremath{\textsc{Positive Planar 1-in-3 SAT}}\xspace}
\newcommand{\sollong}{\ensuremath{\textit{long}}\xspace}
\newcommand{\solmid}{\ensuremath{\textit{midt}}\xspace}
\newcommand{\solshort}{\ensuremath{\textit{short}}\xspace}

@@

Throughout the paper let $ G $ be a \emph{simple graph} with edge set $ S = \{s_1,\dots,s_m\}$ and $ \Gamma $ a straight line drawing of $ G $ in the plane. We call $ \Gamma $ \emph{$ k $-plane} if every edge $ s_i \in S $ is crossed by at most $ k $ other edges from $ S $ in $ \Gamma $. We use the terms edge in $ S $ and segment in $ \Gamma $ interchangeable, since there will always be a drawing $ \Gamma $ given with the graph $ G $.

@@

\section{Hardness of \maxsped for $k\ge3$}\label{sec:hardness}
\fabian{Use the reduction from Till Bruckdorfers Dissertation \cite{b-sgh-15}. The Clauses can be modelled as triangles, the Variables as Circles with every second segment connected to a path. Path should have even length, segments are of length four.}

\soeren{Should the graph in this section too be called G and its drawing $\Gamma$? Also,~\cite{b-sgh-15} claims, paths need to have not only even but equal length, which seems plausible and helps my argument. I'm however not sure if it is actually necessary.}

To show that \maxsped is \NP-hard even for 3-plane input drawings, we will provide a reduction from the \NP-hard \ppsat to \maxsped, similar to Till Bruckdorfers Dissertation \cite{b-sgh-15}. An incremental explanation of \ppsat would be: Given a formula $\phi$ in 3-CNF, in which every single occurring literal is not negated (positive). Consider the graph $P(\phi)$ that contains a vertex for each clause (clause vertex) and a vertex for each variable (variable vertex) and every variable vertex is connected via an edge with a clause vertex, if and only if the variable occurs in the corresponding clause. If $P(\phi)$ is planar, $\phi$ and a planar embedding $\Pi$ of $P(\phi)$ is an instance of \ppsat. The problem is to decide whether $\phi$ has a satisfying truth assignment, such that only one (and since it is a satisfying assignment, exactly one) variable in each clause is set to true.

@@

We create a graph G which is an instance of \maxsped out of an instance of \ppsat by creating the following components.

@@

Let $q$ be the number of clauses and $p$ be the number of distinct variables in $\phi$.

\begin{lemma}
	\label{lem:optimal value}	 
	The construcetd Graph G has an optimal  value in terms of maximizing ink of $8q + 6mp + 9qt - (4q + 2p)\varepsilon$ if $\varepsilon$ is chosen sufficiently small.
\end{lemma}

\begin{proof}
	\soeren{i have an idea for this but i don't know if it would be out of scope, or if i might not even need this lemma. Also is our \maxsped defined as "Does this graph have a drawing with at least x amount of ink?" since thats what this bound is aiming at?"}
\end{proof}
\soeren{it also still needs to be shown that this bound ensures that the two criterias from~\cite{b-sgh-15}, namely that at least one edge in the triangle is drawn and that the segment chains are alternatingly drawn and not drawn.}

\begin{theorem}
	Computing a \maxsped for 3-planar graphs is \NP-hard.
\end{theorem}

\begin{proof}
	Let $\phi$ be a boolean formula and $\Pi$ be a planar embedding corresponding to that formula sucht that ($\phi$, $\Pi$) is an instance of \ppsat. Let $\Gamma$ be a drawing constructed from that instance in the previously described way. We will show that there is a valid truth assignement for $\phi$ if and only if $\Gamma$ admits a \maxsped with a value of at least $8q + 6mp + 3q (3t - 4\varepsilon)$ in terms of maximizing ink.
	~
	\begin{enumerate}
		\item[$\Rightarrow$:] Let $x$ be a variable and $c$ be a clause containing $x$. If $x$ is true draw all red edges in the associated variable gadget completely, which in turn means all green edges are drawn partially and vice versa for when $x$ is false. By drawing the green edges partially we enable the red edges in the segment chains to be drawn completely (if we shorten the stubs for one green edge in the variable gadget by $\varepsilon$). this also means we can draw the edges of the clause gadgets to which this segment chain is connected completely. Since $\phi$ has a satisfying variable assignement where exactly one variable per clause is true we have exactly one of these segment chains which allows exactly one edge per clause graph to be drawn completely. This results in every clause graph having the value $8 - 4\varepsilon$, every variable gadget having the value of $6m - 2\varepsilon$ and every segment chain having the value of $3t$. Let $q$ be the number of clauses and $p$ be the number of distinct variables in $\phi$. This results in

@@

		\item[$\Leftarrow$:]  Assume $\phi$ does not have a stisfying variable assignement. Then there must always exist a clause in which all three contained variables $x_1$, $x_2$ and $x_3$ are false. That however means that, by construction, there is a clause gadget where all the corresponding variable gadgets have all their green edges completely drawn. This adds at most the value of $2\varepsilon$. In order to get the maximal drawing, this propagates through the segment chains up to the clause gadget where it causes all three edges of that gadget to be not drawn. This in turn means that this specific gadget has only the value $6 - 6\varepsilon$. The total value of this drawing in terms of maximizing ink is therefore at least $2$ smaller than normal. If we would try to supplement the missing drawn edge of the clause gadget, we would need to switch up the alternation of drawing segments in the chains, which causes 2 consecutive segments to not be drawn. Since the last edge in the chain was dran before, it now can't be drawn and it therefore again results in a difference of at least $2$ in the complete drawing.
	\end{enumerate}
\end{proof}

@@

>>>>

\newcommand{\ped}{\ensuremath{\textsc{PED}}\xspace}
\newcommand{\maxsped}{\ensuremath{\textsc{MaxSPED}}\xspace}
\newcommand{\ppsat}{\ensuremath{\textsc{Planar 3-Sat}}\xspace}
\newcommand{\sollong}{\ensuremath{\textit{long}}\xspace}
\newcommand{\solmid}{\ensuremath{\textit{midt}}\xspace}
\newcommand{\solshort}{\ensuremath{\textit{short}}\xspace}

@@

Throughout the paper let $ G $ be a \emph{simple graph} with edge set $ S = \{s_1,\dots,s_m\}$ and $ \Gamma $ a straight line drawing of $ G $ in the plane. We call $ \Gamma $ \emph{$ k $-plane} if every edge $ s_i \in S $ is crossed by at most $ k $ other edges from $ S $ in $ \Gamma $. We use the terms edge in $ S $ and segment in $ \Gamma $ interchangeably, since our input is always a graph $G$ with a given drawing $ \Gamma $. Hence we can also interpret $S$ as a set of line segments.

@@

\section{Hardness of \maxsped for $k\ge3$}\label{sec:hardness}
\soeren{Should the graph in this section too be called G and its drawing $\Gamma$? Also,~\cite{b-sgh-15} claims, paths need to have not only even but equal length, which seems plausible and helps my argument. I'm however not sure if it is actually necessary.}

In this section we close the gap between the known hardness of \maxsped~\cite{b-sgh-15} and the polynomial-time algorithm for 2-plane drawings~\cite{bcgkmn-pped-17}.

\begin{theorem}\label{thm:hard}
	The problem \maxsped is \NP-hard even for 3-plane graph drawings.
\end{theorem}

\begin{proof}
	We reduce from the \NP-hard problem \ppsat~\cite{l-pftu-82} using similar ideas as Bruckdorfer's general proof of the hardness of \maxsped~\cite{b-sgh-15}. 
	Let $\phi$ be a planar 3-Sat formula with $n$ variables $\{x_1, \dots, x_n\}$ and $m$ clauses $\{c_1, \dots, c_m\}$, each consisting of three literals.
	We can assume that $\phi$ comes with a planar drawing of its variable-clause graph $H_\phi$ that has a vertex for each variable $x_i$ placed on a horizontal line and a vertex for each clause $c_j$ connected to the three variables appearing in $c_j$. 
	In our gadget-based reduction we create a 3-plane drawing $\Gamma_\phi$ as a set of line segments of uniform length and a value $L$ such that $\Gamma$ has a SPED of total ink at least $L$ if and only if $\phi$ is satisfiable.

	All segments

% To show that \maxsped is \NP-hard even for 3-plane input drawings, we will provide a reduction from the \NP-hard \ppsat to \maxsped, similar to Till Bruckdorfers Dissertation \cite{b-sgh-15}. An incremental explanation of \ppsat would be: Given a formula $\phi$ in 3-CNF, in which every single occurring literal is not negated (positive). Consider the graph $P(\phi)$ that contains a vertex for each clause (clause vertex) and a vertex for each variable (variable vertex) and every variable vertex is connected via an edge with a clause vertex, if and only if the variable occurs in the corresponding clause. If $P(\phi)$ is planar, $\phi$ and a planar embedding $\Pi$ of $P(\phi)$ is an instance of \ppsat. The problem is to decide whether $\phi$ has a satisfying truth assignment, such that only one (and since it is a satisfying assignment, exactly one) variable in each clause is set to true.

@@

% We create a graph G which is an instance of \maxsped out of an instance of \ppsat by creating the following components.

@@

% Let $q$ be the number of clauses and $p$ be the number of distinct variables in $\phi$.
%
% \begin{lemma}
% 	\label{lem:optimal value}
% 	The construcetd Graph G has an optimal  value in terms of maximizing ink of $8q + 6mp + 9qt - (4q + 2p)\varepsilon$ if $\varepsilon$ is chosen sufficiently small.
% \end{lemma}
%
% \begin{proof}
% 	\soeren{i have an idea for this but i don't know if it would be out of scope, or if i might not even need this lemma. Also is our \maxsped defined as "Does this graph have a drawing with at least x amount of ink?" since thats what this bound is aiming at?"}
% \end{proof}
\soeren{it also still needs to be shown that this bound ensures that the two criterias from~\cite{b-sgh-15}, namely that at least one edge in the triangle is drawn and that the segment chains are alternatingly drawn and not drawn.}

	Let $\phi$ be a boolean formula and $\Pi$ be a planar embedding corresponding to that formula sucht that ($\phi$, $\Pi$) is an instance of \ppsat. Let $\Gamma$ be a drawing constructed from that instance in the previously described way. We will show that there is a valid truth assignement for $\phi$ if and only if $\Gamma$ admits a \maxsped with a value of at least $8q + 6mp + 3q (3t - 4\varepsilon)$ in terms of maximizing ink.
	~
	\begin{enumerate}
		\item[$\Rightarrow$:] Let $x$ be a variable and $c$ be a clause containing $x$. If $x$ is true draw all red edges in the associated variable gadget completely, which in turn means all green edges are drawn partially and vice versa for when $x$ is false. By drawing the green edges partially we enable the red edges in the segment chains to be drawn completely (if we shorten the stubs for one green edge in the variable gadget by $\varepsilon$). this also means we can draw the edges of the clause gadgets to which this segment chain is connected completely. Since $\phi$ has a satisfying variable assignement where exactly one variable per clause is true we have exactly one of these segment chains which allows exactly one edge per clause graph to be drawn completely. This results in every clause graph having the value $8 - 4\varepsilon$, every variable gadget having the value of $6m - 2\varepsilon$ and every segment chain having the value of $3t$. Let $q$ be the number of clauses and $p$ be the number of distinct variables in $\phi$. This results in

@@

		\item[$\Leftarrow$:]  Assume $\phi$ does not have a stisfying variable assignement. Then there must always exist a clause in which all three contained variables $x_1$, $x_2$ and $x_3$ are false. That however means that, by construction, there is a clause gadget where all the corresponding variable gadgets have all their green edges completely drawn. This adds at most the value of $2\varepsilon$. In order to get the maximal drawing, this propagates through the segment chains up to the clause gadget where it causes all three edges of that gadget to be not drawn. This in turn means that this specific gadget has only the value $6 - 6\varepsilon$. The total value of this drawing in terms of maximizing ink is therefore at least $2$ smaller than normal. If we would try to supplement the missing drawn edge of the clause gadget, we would need to switch up the alternation of drawing segments in the chains, which causes 2 consecutive segments to not be drawn. Since the last edge in the chain was dran before, it now can't be drawn and it therefore again results in a difference of at least $2$ in the complete drawing.
	\end{enumerate}

@@

\end{proof}
